---
title: "Lab 6"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# R Lab 6
# Exercise 1
## 1.1
- X is V2, which means the ACT test score.
- Y is V1, which means the GPA at the end of the freshman year.
```{r}
GPA = read.table("./data/CH01PR19.txt")
attach(GPA)
X <- V2
Y <- V1
# X = ACT, Y = GPA
```
## 1.2
According to the plot below, we can see that many data points do not fit the linear regression line. 
```{r}
reg = lm(Y ~ X)
reg
plot(X, Y, xlab = "x = ACT", ylab = "y = GPA")
abline(reg, col = "red")
# b0 =  2.11405   
# b1 = 0.03883  
```

## 1.3
We use `predict` function to estimate it. If ACT score is 30, then the estimated mean of the freshman GPA would be 3.278863.
```{r}
predict(reg, data.frame(X = 30))
```
## 1.4
The key is we should find b1. If the ACT score increases by one point, we estimate GPA would be changed 0.03882713 units.
```{r}
summary(reg)
reg$coefficients[2]
```
## 1.5
e is the residuals. Another way is we can use `anova` table to find out the residuals.
```{r}
e = Y - fitted.values(reg)
e # ei each
sum(e^2) # sum of the squared residuals 

anova(reg)
```
## 1.6
estimated variance is 0.3882848, and residual standard error: 0.6231 on 118 degrees of freedom
```{r}
n <- length(X)
var_est <- sum(e^2)/(n-2) # sum of the squared residuals - degrees of freedom (n -2, b0 and b1 so minus 2)
var_est
sqrt(var_est)
summary(reg)
```

# Exercise 2
## 2.1
Zero is not included, because the b1 is 0.005385614 to 0.07226864.
```{r}
confint(reg, level = 0.99)
```
## 2.2
- Ho: $\beta1 = 0$ (have NOT a linear association between ACT score (X) and GPA at the end of the freshman year (Y)).
- Ha: $\beta1 \neq 0$ (have a linear association between ACT score (X) and GPA at the end of the freshman year (Y)).

According to tables of a linear regression model above, we have evidence to reject the null hypothesis in favor of the alternative hypothesis with the 99 % confidence interval. Thus, there is a linear association between student's ACT score (X) and GPA at the end of the freshman year (Y).

## 2.3
For the actual **population mean** response - confidence intervals\

If ACT score is 28, we have 95% certain contains the population mean of freshman GPA is 3.061384 to 3.341033.
```{r}
predict(reg, data.frame(X = 28), interval = "confidence", level = 0.95)
```

## 2.4
For the individual response (actual response) - prediction intervals\

If student obtained a 28 on the ACT, the 95% prediction interval is 1.959355 to 4.443063 of his freshman GPA.
```{r}
predict(reg, data.frame(X = 28), interval = "prediction")
```

## 2.5
The majority of data points are in the range of upper band and lower band. We can conclude that the true regression relation has been precisely estimated.
```{r}
n = length(X) #sample sizes
e = reg$residuals # residuals
s = sqrt(sum(e^2)/(n-2)) # estimated standard deviation = root MSE
s
W = sqrt(qf(0.95,2,n-2))  # quantity of F-distribution
W
Yhat = fitted.values(reg) # Yhat = b0 + b1x = predict(reg)
Sxx = (n-1)*var(X) 

margin = W*s*sqrt(1/n + (X - mean(X))^2/Sxx) 
upper.band = Yhat + W*s*sqrt(1 + 1/n + (X - mean(X))^2/Sxx) 
lower.band = Yhat - W*s*sqrt(1 + 1/n + (X - mean(X))^2/Sxx) 

plot(X,Y,xlab="ACT", ylab="Y = GPA", xlim = c(20,30))
abline(reg,col="red")
lines(X,upper.band,col="blue")
lines(X,lower.band,col="blue")
```

