---
title: "Lab 13"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output: pdf_document
urlcolor: blue
linkcolor: red
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(leaps)
```

# Fitting a polynomial regression
Because the distribution of observed data does not match the linear regression line at the beginning. We decide not to use the linear regression model.
```{r}
library(ISLR)
attach(Auto)

reg <- lm(mpg~weight, data = Auto)

Auto %>%
  ggplot(aes(weight, mpg))+
  geom_smooth(method = lm, se = FALSE) +
  geom_point() +
  theme_bw()
```

# look at the residual plot
Examine the pattern in the residuals vs weight plot, which indicates a bad bit (but not really bad). To achieve this goal, consider using a more reliable regression model.
```{r}
# plot( weight, residuals(reg))
augment(reg) %>%
  ggplot(aes(x = weight, y = .resid)) +
  geom_smooth(method = lm, se = FALSE) +
  geom_point() +
  theme_bw()
```

# summary Table
```{r}
summary(reg)
```

# Try to use polynomial regression model
## model selection
```{r}
polynomial.fit <-  regsubsets(mpg ~ poly(weight, 10), data=Auto )
summary(polynomial.fit)
```

# find out the largest adjusted R squares

```{r}
summary(polynomial.fit)$adjr2
which.max(summary(polynomial.fit)$adjr2)
```
So the largest adjusted R squares is 
$$
\hat{y} = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + \beta_4X^4 + e
$$
# find out the smallest CP
```{r}
summary(polynomial.fit)$cp
which.min(summary(polynomial.fit)$cp)
```
So the smallest CP is 
$$
\hat{y} = \beta_0 + \beta_1X + \beta_2X^2 + \beta_3X^3 + e
$$
# find out the smallest BIC (penalized-likelihood criteria)
```{r}
summary(polynomial.fit)$bic
which.min(summary(polynomial.fit)$bic)
```

so the smallest BIC is
$$
\hat{y} = \beta_0 + \beta_1X + \beta_2X^2  + e
$$

# Cross-validation agrees with the quadratic model
Assessing how the results of a statistical analysis will generalize to an independent data set. 
```{r}
library(boot)
cv.error = rep(0,10)
for (p in 1:10){
  polynomial = glm( mpg ~ poly(weight, p) )
  cv.error[p] = cv.glm( Auto, polynomial )$delta[1]}

cv.error
```

# Plot the cv.error
- The `y` index represents the **Mean squared error**. Our goal is to find the polynomial model with the lowest MSE. The quadratic model has the lowest MSE, according to the results.
- So, we choose the quadratic regression – degree 2 polynomial. Its prediction MSE is 17.52474.
```{r}
plot(cv.error); lines(cv.error)
which.min(cv.error)
```

```{r}
poly2 <- lm( mpg ~ poly(weight,2), data = Auto)
summary(poly2)
```

# Plot
```{r}
Auto %>%
  ggplot(aes(x = weight, y = mpg)) +
  stat_smooth(method="lm", se=TRUE, fill=NA,
                formula= y ~ poly(x, 2, raw = TRUE), colour="red") +
  geom_point()
```

The quadratic regression best fits the data in this lab. We can conclude that this model is good because it has a higher Adjusted R-squared (0.7137) and each independent variable is significant.


# Reference
- https://stats.stackexchange.com/questions/95939/how-to-interpret-coefficients-from-a-polynomial-model-fit
