---
title: "COVID-19 Deaths Determinant"
author: "Yunting Chiu, Ana Hyo Young Lim, ChiYun Liu"
date: "`r Sys.Date()"
output:
  html_document:
    number_sections: yes
    theme: cerulean
  pdf_document: default
urlcolor: blue
linkcolor: red
---

```{r include=FALSE}
library(tidyverse)
library(leaps)
library(car)
library(MASS)
```

## Covid Data Input
```{r}
covid_dftmp <- read_csv("../data/covid_df_regression.csv")

covid_dftmp %>% 
  dplyr::select(-Province_State, -Total_Population) -> covid_df
```

## Variable selection

1. Full model

```{r}
attach(covid_df)
reg_covid_df <- lm(deaths ~., data = covid_df)
summary(reg_covid_df)
```

2. Compute best subsets regression
```{r message=FALSE}
reg_fit <- regsubsets(deaths ~., nvmax = 10, data =  covid_df)
summary(reg_fit)
```

Summary table above finds the best model for each set of variables, but we have to choose one best set of variable for the model for analyzing.

3. Select variable and choose the optimal model

```{r}
summary(reg_fit)$adjr2 # find max adjr2---> p = 6,0.9265559
summary(reg_fit)$cp # find min cp ---> p = p = 4,3.257537
summary(reg_fit)$bic # find min bic ---> p = 4, -117.23954

which.max(summary(reg_fit)$adjr2)
which.min(summary(reg_fit)$cp) 
which.min(summary(reg_fit)$bic)
```

* Adjusted $R^2$ reports the best model is the one with 6 variables(confirmed + Asian_alone +  Black_or_African_American_alone + Some_other_race_alone + White_alone + IncomeTwenty)

* **BIC and Mallows** $C_p$ **criteria** give the best model with 4 variables(confirmed + Asian_alone + White_alone + IncomeTwenty).

4. Use backward elimination variable selection

```{r}
# backward elimination variable selection
reg_fit_bk <- regsubsets(deaths ~., data = covid_df, method = "backward")
plot(reg_fit_bk, scale = "adjr2")
plot(reg_fit_bk, scale = "bic")
```

* backward elimination variable selection:

**BIC** reports the best mode should be included with confirmed + Asian_alone + white_alone + IncomeTwenty

**adjr2** reports the best mode should be included with confirmed + Asian_alone + Black_or_African_American_alone + other_race_alone + white_alone + IncomeTwenty


5. Use a stepwise procedure to choose the best model

```{r}
null = lm(deaths ~ 1, data = covid_df)
full = lm(deaths ~., data = covid_df)
step(null, scope = list(lower = null, upper = full), direction = "forward")
```

 * **AIC **reports the best model with 5 variables is deaths ~ White_alone + Black_or_African_American_alone + IncomeTwenty + Asian_alone + confirmed
 

## test model

1. BIC/AIC model and summary
```{r}
# BIC---4
reg_model_bic <- lm(deaths ~ confirmed + Asian_alone + White_alone + IncomeTwenty, data = covid_df)
summary(reg_model_bic)

# AIC---5
reg_model_aic <- lm(deaths ~ White_alone + Black_or_African_American_alone + IncomeTwenty + Asian_alone + confirmed, data = covid_df)
summary(reg_model_aic)

## seems AIC better
```


2. Diagnose model AIC
```{r}
par(mfrow = c(2,2))
plot(reg_model_aic)
```

```{r}
t_aic <- rstudent(reg_model_aic)
shapiro.test(t_aic) 
```

* p < $\alpha$, not normal distributed


```{r}
ncvTest(reg_model_aic)
```

* Homo is rejected


3. Fixes non-normality
```{r}
bc_aic <- boxcox(reg_model_aic)
lambda <- bc_aic$x[which.max(bc_aic$y)]
lambda
```
```{r}
# test -2
reg_model_aic1 <- lm(deaths^(-2) ~ White_alone + Black_or_African_American_alone + IncomeTwenty + Asian_alone + confirmed, data = covid_df)
new_t_aic1 <- rstudent(reg_model_aic1)
shapiro.test(new_t_aic1)
```


```{r}
# test best lambda
reg_model_aic2 <- lm(deaths^lambda ~ White_alone + Black_or_African_American_alone + IncomeTwenty + Asian_alone + confirmed, data = covid_df)
new_t_aic2 <- rstudent(reg_model_aic2)
shapiro.test(new_t_aic2)
```

```{r}
par(mfrow = c(2,2))
plot(reg_model_aic2) ## normal qq pass normality
```
```{r}
step(reg_model_aic2) 
ncvTest(reg_model_aic2) # does not pass constant variance, conclude heter
outlierTest(reg_model_aic2) # pass bonferroni p

# lack of fit
# reduced = lm(deaths^lambda ~ White_alone + Black_or_African_American_alone + IncomeTwenty + Asian_alone + confirmed, data = covid_df1)
# full = lm(deaths^lambda ~ as.factor(White_alone) + as.factor(Black_or_African_American_alone) + as.factor(IncomeTwenty) + as.factor(Asian_alone) + as.factor(confirmed), data = covid_df1)
# anova(reduced, full)
```

4. Remove outliers and test model again
```{r}
covid_df1 <- covid_df[-c(23,31,10),]
reg_model_aic_rmout <- lm(deaths^lambda ~ White_alone + Black_or_African_American_alone + IncomeTwenty + Asian_alone + confirmed, data = covid_df1)
ncvTest(reg_model_aic_rmout)  # pass constant variance, conclude homo

new_t_aic_rmout <- rstudent(reg_model_aic_rmout)
shapiro.test(new_t_aic_rmout)
```
5. AIC = 415.82, choose new model again

```{r}
reg_model_aic3 <- lm(deaths^lambda ~ White_alone + IncomeTwenty + Asian_alone + confirmed, data = covid_df)
new_t_aic3 <- rstudent(reg_model_aic3)
shapiro.test(new_t_aic3)
```
```{r}
par(mfrow = c(2,2))
plot(reg_model_aic3)
```



3. Diagnose model BIC
```{r}
par(mfrow = c(2,2))
plot(reg_model_bic)
```
```{r}
# covid_df1 <- covid_df[-c(5,10,33),]
# BIC
reg_model_bic1 <- lm(deaths ~ confirmed + Asian_alone + White_alone + IncomeTwenty, data = covid_df1)
summary(reg_model_bic1)

par(mfrow = c(2,2))
plot(reg_model_bic1)
t_bic1 <- rstudent(reg_model_bic1)
shapiro.test(t_bic1) 
```


```{r}
# BIC
t_bic <- rstudent(reg_model_bic)
shapiro.test(t_bic) 
```
* p < $\alpha$, not normal distributed

```{r}
boxcox(reg_model_bic)
```


```{r}
ncvTest(reg_model_bic)
```

* Homo is rejected





