---
title: "Homework #2"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output: pdf_document
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. (**1.2**) The members of a health spa pay annual membership dues of $\$$300 plus a charge of $\$$2 for each visit to the spa. Let Y denote the dollar cost for the year for a member and X the number of visits by the member during the year. Express the relation between X and Y mathematically. Is it a functional relation or a statistical relation (that is, is the relation deterministic or stochastic)?

- deterministic: the output of the model is entirely determined by the values of the parameters and the initial conditions; stochastic: random, unpredictable.
- The association between X and Y is: Y = 300 + 2X (dollars)
This is a functional relationship because a effect in the value of the X will cause the corresponding change in the value of the Y. X and Y will not have a uncertain effect.

2. (**1.6**) Suppose the regression parameters are $\beta0$ = 200 and $\beta1$ = 5.0.\
(a) Plot the regression equation.\
(b) Predict the response for X = 10, 20, and 40.\
(c) Explain the meaning of parameters $\beta0$ and $\beta1$.\

3. (**1.10**) An analyst in a large corporation studied the relation between current annual salary (Y ) and age (X) for the 46 computer programmers presently employed in the company. The analyst concluded that the relation is curvilinear, reaching a maximum at 47 years. Does this imply that the salary for a programmer increases until age 47 and then decreases? Explain.

```{r}
library(tidyverse)
aaa <- read_tsv("./data/CH01PR19.txt")
aaa
```

