---
title: "Matrix Operations and Multivariate Linear Regression"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# R Lab 9: Matrix Operations and Multivariate Linear Regression

Define a matrix by entering its elements manually.
```{r}
B <-  matrix(c(1,2,3,4,5,6,7,8,9),3,3)
B
```
This is an element-wise operation
```{r}
B^2
```

This is matrix multiplication, $B^2$ = B*B
```{r}
B %*% B
```

Transposed matrix
```{r}
t(B)
```


Joining two matrices side by side (as columns)
```{r}
cbind(B, B)
```

Joining two matrices below each other (as rows)
```{r}
rbind(B, B)
```

Sub-matrix, a part of matrix B
```{r}
B[1:3, 1:2]
```

Inverting matrices is available in package “matlib”
```{r}
# install.packages("matlib")
library(matlib)
```

```{r, error=TRUE}
B
inv(B)
```
This means there is a linear dependence among columns (and among rows) of matrix B. Such matrices are not invertible, and they have a determinant equal det(B)=0

make determinant equal det(B)=0
```{r}
det(B)
```

```{r}
B[1,1]=100
B
```

We changed the matrix by adding a “ridge”, and now the inverse B^-1 exists
```{r}
inv(B)
```

# Data practice - Multivariate Linear Regression
Define a matrix from the “mtcars” data set and build a regression model that **predicts miles per gallon** based on the number of cylinders, horsepower, axes ratio, weight, and acceleration time.
```{r}
head(mtcars)
```
Extract X-matrix of predictors
```{r}
X = data.matrix(mtcars[,c(2,4:7)])
head(X)
```
Find the vector of responses of Y 
```{r}
Y = data.matrix(mtcars[ ,1])
head(Y)
```

**We also need a vector of 1s to include the intercept**
```{r}
n = length(Y)
one = matrix(1,n,1) #row 1, col n, values 1
n
one
```

```{r}
X = cbind(one,X)
X
```

This is matrix X’X
```{r}
t(X) %*% X 
```

Slope $\beta1$ = (X’X)-1X’Y
```{r}
slope = inv(t(X) %*% X) %*% t(X) %*% Y
slope
```

# Another method to get thesame slope by using built-in function `lm`
```{r}
attach(mtcars)
recheckReg <- lm(mpg ~ cyl + hp + drat + wt + qsec)
recheckReg
```

Our estimated regression equation is `mpg = 25.95 – 0.49 cyl – 0.015 hp + 1.13 drat – 3.38 wt + 0.35 qsec + e`

