---
title: 'COVID19 Confirmed vs Deaths'
subtitle: "Web Scraping Data from Johns Hopkins University"
author: "Yunting Chiu"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    theme: cerulean
  pdf_document: default
urlcolor: blue
linkcolor: red
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align  = "center",
                      fig.height = 5, 
                      fig.width  = 6)
```


# Web Scrapping Covid Data
- The data source is from [Johns Hopkins University](https://github.com/CSSEGISandData)
- "https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series"
- If you instead download the file using the github.com link, you will actually be downloading a web page with buttons and comments and which displays your desired content in the middle. It's what you want to give to your web browser to get a nice page to look at but not to download.
```{r}
# Installed library
library(tidyverse)
url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"

```

# Extract four datasets
    + time_series_covid19_confirmed_global.csv
    + time_series_covid19_deaths_global.csv
    + time_series_covid19_confirmed_US.csv
    + time_series_covid19_deaths_US.csv
```{r}
df <- tibble(file_names = c("time_series_covid19_confirmed_global.csv",
                            "time_series_covid19_deaths_global.csv",
                            "time_series_covid19_confirmed_US.csv",
                            "time_series_covid19_deaths_US.csv")) -> df
```

# Mapping Data
```{r}
df %>%
  mutate(url = str_c(url_in, file_names, sep = "")) -> df
```
```{r}
df %>%
  mutate(data = map(url, ~read_csv(., na = ""))) -> df
```
```{r}
df %>%
  mutate(case_types = as.factor(str_extract(file_names, "[:alpha:]*_[gU][:alpha:]*"))) -> 
  df
# alpha = Any letter, [A-Za-z]
# reference: https://www.petefreitag.com/cheatsheets/regex/character-classes/
```
```{r}
df %>%
  dplyr::select(case_types, data) -> df
```

# Clean Data  
```{r}
df %>%
  mutate(vars = map(df$data, names)) -> df
# map(df$vars, ~unlist(.)[1:15]) for checking

fix_names <- function(df, pattern, rePattern){
  stopifnot(is.data.frame(df), is.character(pattern), is.character(rePattern))
  names(df) <- str_replace_all(names(df), pattern, rePattern)
  return(df)
}

df %>%
  mutate(data = map(data, ~fix_names(., "([ey])/", "\\1_")),
         data = map(data, ~fix_names(., "Admin2", "County")),
         data = map(data, ~fix_names(., "Long_", "Long")),
         data = map_if(data, str_detect(df$case_types, "US"),
                   ~select(., -c("UID", "iso2", "iso3", 
                                 "code3", "FIPS", "Combined_Key"))),
         data = map_if(data, str_detect(df$case_types, "global"),
                      ~mutate(., County = "NA")),
         data = map_if(data, !str_detect(df$case_types, "deaths_US"),
                      ~mutate(., Population = 0)),
         data = map(data, ~unite(., "Country_State", 
                                 c("Country_Region", "Province_State"),
                                 remove = FALSE, na.rm = TRUE,
                                 sep = "_"))
         ) -> df

df %>%
  mutate(vars = map(df$data, names)) -> df # synchronize the vars correspondingly
# map(df$vars, ~unlist(.)) # for checking 
```

# Tidy each dataframe 

```{r}
library(lubridate)
df %>%
  mutate(data = map(data, ~pivot_longer(data = ., cols = contains("/"),
                                        names_to = "Date",
                                        values_to = "dailyValues",
                                        names_transform = list(Date = mdy)))
         ) -> df
# df$data <- map(df$data, names) # synchronize the vars correspondingly
# map(df$vars, ~unlist(.)) # for checking 

# crate a function to fix in type of Date
mdyDate <- function(df, varsDate){
  # stopifnot(is.data.frame(df), is.character(varsDate))
  df[[varsDate]] <- ymd(df[[varsDate]])
  return(df)
}

df %>%
  mutate(data = map(data, ~mdyDate(., "Date"))) -> df_long

# str(df_long) # check the data set
```

# Add Continents 
```{r, warning = FALSE}
library(countrycode) # get the continents
df_long %>%
  mutate(data = map(data, ~mutate(., Continent = countrycode(Country_Region,
                                               origin = "country.name",
                                               destination = "continent")))
         ) -> df_long
```

# Fix NAs for Continents
```{r}
df_long %>%
  mutate(data = map(data, ~mutate(., Continent = case_when(
                                               Country_Region == "Diamond Princess" ~ "Asia",
                                               Country_Region == "Kosovo" ~ "Americas",
                                               Country_Region == "MS Zaandam" ~ "Europe",
                                               TRUE ~ Continent)
                                  ))) -> df_long

map(df_long$data, ~unique(.$Continent))
```

# Unnest the Data Frames    
```{r}
# 1
df_long %>%
  unnest(cols = data) %>%
  ungroup() -> df_all

# 2
remove(df, df_long)

# 3
df_all %>%
  select(-vars) -> df_all
```

# Get World Population Data
- source: [UN source](https://population.un.org/wpp/Download/Standard/CSV/)
```{r}
# 1
df_pop <- read_csv("./data/WPP2019_TotalPopulation.csv")
# summarize(df_pop, across(everything(), ~sum(is.na(.)))) # check NAs

# 2 
semi_join(df_pop, df_all, by = c("Location" = "Country_Region")) -> df_pop

# 3
df_pop %>% 
  mutate(rank_p = rank(-PopTotal, na.last = TRUE),
         rank_d = rank(-PopDensity, na.last = TRUE),
         PopTotal = (PopTotal*1000)) -> df_pop
```

# Add Population Data to `df_all`
```{r}
df_all %>%
  inner_join(df_pop, by = c("Country_Region" = "Location")) -> df_all

df_all 
```

# We only focus on 2020/01/22 - 2021/01/22 
```{r}
# extract one year
df_all %>%
  filter(case_types == "confirmed_US" & as_date(Date) <= as_date("2021-01-22") | case_types == "deaths_US" & as_date(Date) <= as_date("2021-01-22")) -> covid
```

# find out each US state using usmap
```{r}
library(usmap)
state_map <- us_map(regions = "states")
state_map %>%
  distinct(full) %>%
  rename("Province_State" = "full") -> USstates
```
# Obtain the number of confirmed cases for each state
```{r}
covid %>%
  filter(case_types == "confirmed_US" & as_date(Date) == as_date("2021-01-22")) %>%
  select(Province_State, County, dailyValues) %>%
  group_by(Province_State) %>%
  tally(dailyValues) %>%
  right_join(USstates) %>%
  rename("confirmed" = "n") -> confirmed
```
# Obtain the number of death cases for each state
```{r}
covid %>%
  filter(case_types == "deaths_US" & as_date(Date) == as_date("2021-01-22")) %>%
  select(Province_State, County, dailyValues) %>%
  group_by(Province_State) %>%
  tally(dailyValues) %>%
  right_join(USstates) %>%
  rename("deaths" = "n") -> deathes

full_join(confirmed, deathes) -> covidForRegression
```

# Example of simple linear regression model
```{r}
reg <- lm(deaths ~ confirmed, data = covidForRegression)
summary(reg)

covidForRegression %>%
  ggplot(aes(x = confirmed, y = deaths)) +
  geom_smooth(method = lm) +
  geom_point()
```

# save the dataframe to a local drive
```{r}
#covid %>%
  #write_excel_csv("covidJHU.csv")
```
